Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-04-13 ltx/BrainIDE.tex

\section{State and Parameter Estimation}
The Unscented Rauch Tung Striebel Smoother (URTSS) was used for the state estimation \cite{Sarkka2010}, where the Unscented Kalman filter is used for the forward iteration \cite{Julier1997}. This method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to this approach are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0=\mathbb E[\mathbf x_0]
\end{equation*}
\begin{equation*}
\mathbf P_0=\mathbb E[(\mathbf x_0-\hat{\mathbf x}_0)(\mathbf x_0-\hat{\mathbf x}_0)^\top]
\end{equation*}
\State Form the sigma vectors 
\begin{equation*}
\mathcal X_k=[\hat{\mathbf x}_k \quad \hat{\mathbf x}_k+\gamma\sqrt{\mathbf P_k} \quad \hat{\mathbf x}_k-\gamma\sqrt{\mathbf P_k}]
\end{equation*}
\State Update the sigma vectors by propogating them through state transition equation and compute the predicted state and the predicted state covariance
\begin{equation*}
 \hat{\mathbf x}_{k +1|k}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,k+1|k}
\end{equation*}
\begin{equation*}
 \mathbf P_{k +1|k}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,k+1|k}-\hat{\mathbf x}_{k +1|k})(\mathcal X_{i,k+1|k}-\hat{\mathbf x}_{k +1|k})^\top+\boldsymbol \Sigma_w
\end{equation*}
\State use extrema to form the splines $s_{min,k}(n)$ and $s_{max,k}(n)$
\State calculate $x_{0,k}(n) = 0.5\left(s_{min,k}(n) + s_{max,k}(n)\right)$
\State find local maxima of $h_k(n)$
\State use maxima to form the spline $r_k(n) = s_{max,k}(n)$
\While { $x_{0,k}(n)$ is not monotonic }
    \State $k=k+1$
    \State perform CPT on $x_{0,k-1}(n)$ to estimate $\phi_{k}(n)$
    \State repeats steps 3 to 7 to approximate parameters
\EndWhile

\end{algorithmic}
\end{algorithm}
\section{Comparison of Estimated Parameters and True Parameters}

>>>>

\section{State and Parameter Estimation}
The Unscented Rauch Tung Striebel Smoother (URTSS) was used for the state estimation \cite{Sarkka2010}, where the Unscented Kalman filter is used for the forward iteration \cite{Julier1997}. This method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to this approach are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation with at least second order accuracy. In addition, the computation of the Jacobian used in the EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unscented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filtering step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm

\section{Results}
\subsection{Simulation}
Figure~\ref{} illustrates the simulation used to generate the data. Part A. of the Figure shows the simulated field with the sensors positioned location

\subsection{Frequency Analysis}
Figure~\ref{SprectralAnalysis} shows the spectra of the connectivity kernel, the disturbance covariance, the field filtered by the sensor kernel and the observations.

\subsection{State and Parameter Estimation}
